---
title: "604 Group MLR"
author: "Jianling Xie, Alan Cheun, Zane Wu"
date: "2024-03-27"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

Modelling Plan We will first run a First order linear regression model
using all predictors and test the variables for multicolinearity. Once
we have removed the variables with high multicolinearity, we will use
step-wise regression to select a model of main effects. We will then
perform a partial F-test to compare the full model and reduced model.
Once we decide the main effects, we will use the individual t-test to
check for significant higher-order terms and interactions. Then we test
this model with another F-test to evaluate if the higher order terms and
interactions are significant. Any significant higher-order or
interaction terms will be added to the main effects to produce the
finalnal model. Our model will then test for the following 6 assumptions
as shown below:

1.  Linearity Assumption - Review residual plots
2.  Independence Assumption - Review residual against year (time)
3.  Normality Assumption - Using Shapiro-Wilk normality test
4.  Equal Variance Assumption (heteroscedasticity) - Using Breusch-Pagan
    test
5.  Multicollinearity - Using variance infation factors (VIF)
6.  Outliers - check Cook's distance and leverage

Finally, we may use the final model to do a prediction of crime at a
specific future c-train station

dependent variable: crime_rate ln_crime_rate Total_Crime_Count

independent variables: Year Sectors SHORTEST_DISTANCE_TO_LRT_METERS
SHORTEST_DISTANCE_TO_POLICE_METERS male_percentage_num
age_75_plus_percentage_num TotalPermits
calgary_cma_average_hourly_wage_rate_num median
canada_unemployment_rate_num

```{r}

library(car)
library(lmtest)
library(olsrr)
library(Ecdat)
library(MASS)
library(dplyr)
library(knitr)
library(kableExtra)
library(mctest)
library(pastecs)
library(psych)

```

# Dataset

Our dataset is sourced from The City of Calgary’s Open Data Portal. It
is the outcome of our DATA604 project to analyze Calgary Crime
Statistics.

License:
<https://data.calgary.ca/stories/s/Open-Calgary-Terms-of-Use/u45n-7awa>

```{r}

data <- read.csv("https://raw.githubusercontent.com/ANPC86/D603/main/603%20crime%20dataset%20April%201%202024.csv")
str(data)



```

```{r}
data$CLASS_CODE.f = as.factor(data$CLASS_CODE)

## create dependent variables: crime rate and log of crime rate
data<- data %>%
  mutate(
    crime_rate = Total_Crime_Count / Avg_Resident_Count * 100000,
    ln_crime_rate = log(crime_rate)
  )

```

```{r}
table (data$CLASS_CODE.f) # cell count for level-2 (industry) was low, it may cause problem, so decided not to include it to the model

table (data$Sectors) # OK

```

## Check Dataset for Missing Values

Check for missing values - 25\~30 with missing data

```{r}
missing_values <- colSums(is.na(data))

missing_values

```

Using complete case analysis

```{r}

# Select variables and create a new dataset without missing data

data_selected <- data[, c("Year",
                          "crime_rate",
                          "ln_crime_rate",
                          "Total_Crime_Count", 
                          "Avg_Resident_Count",
                          "Sectors", 
                          "SHORTEST_DISTANCE_TO_LRT_METERS", 
                           "SHORTEST_DISTANCE_TO_POLICE_METERS", "male_percentage_num", 
                           "age_75_plus_percentage_num", "TotalPermits", 
                           "calgary_cma_average_hourly_wage_rate_num", 
                           "median", 
                           "CLASS_CODE.f",
                          "canada_unemployment_rate_num")]

# Remove rows with missing values
data_complete <- na.omit(data_selected)

# Print the new dataset
str(data_complete)
```

## Test Distribution of the Dependent Variable - Crime Rate

Check the distribution of the dependent variable crime rate. The crime
rate was highly right-skewed, and applying a log transformation helped
approximate a normal distribution.

```{r}

par(mfrow = c(1, 2))
hist(data_complete$crime_rate, main='Crime rate', xlab='rate/100,000')
hist(data_complete$ln_crime_rate,main='LN (crime rate)', xlab='rate/100,000') 
par(mfrow = c(1, 1))

describe(data_complete$ln_crime_rate)

```

# Fit the Additive Model 1

Fit the full model with all candidate variables, and check for
multicollinearity

```{r}

model1 <- lm(ln_crime_rate ~ 
Year+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
calgary_cma_average_hourly_wage_rate_num+
median+
canada_unemployment_rate_num, data_complete)

summary(model1)


```

## Check Model 1 for MultiCollinearity

```{r}
imcdiag(model1, method="VIF")
```

The variable "calgary_cma_average_hourly_wage_rate_num" has
VIF=37.604408, p = 0.00479 The variable "Year" had VIF=38.994258,
p=5.15e-05. So, we dropped variable
"calgary_cma_average_hourly_wage_rate_num", re-run the reduced model,
and test for multicollinearity again

# Fit Reduced Model 2

```{r}

model2 <- lm(ln_crime_rate ~ 
Year+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete)

summary(model2)


```

## Check Model 2 for MultiCollinearity

```{r}
imcdiag(mod = model2, method = "VIF")

```

In model2, all VIF's \< 5.0, suggesting moderate collinearity, but it is
not severe enough to warrant corrective measures.

Although we planned to use step-wise for selecting variables, but it is
not necessary at this point, as in model2, the individual t-test for all
the candidate variables had p\<0.05.

# Fit Interaction Model 3

So we move on the testing all possible two way interactions:

```{r}

model3 <- lm(ln_crime_rate ~ 
(Year+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num)^2, data_complete)

summary(model3)


```

Evaluate interaction terms significance using individual t-test.

$$
H_0:\\
H_a:\\
\alpha=0.05
$$

The following interaction terms had p\<0.05:

Year:male_percentage_num 2.413 0.015997 \*\
Year:TotalPermits 5.620 2.41e-08 ***Year:canada_unemployment_rate_num
-2.091 0.036737*** **\
SectorsEAST:SHORTEST_DISTANCE_TO_LRT_METERS -3.098 0.001995**
SectorsNORTH:SHORTEST_DISTANCE_TO_LRT_METERS 1.966 0.049571 \*\
SectorsSOUTH:SHORTEST_DISTANCE_TO_LRT_METERS -2.853 0.004406 \*\*
SectorsSOUTHEAST:SHORTEST_DISTANCE_TO_LRT_METERS 2.134 0.033069 \*\
SectorsWEST:SHORTEST_DISTANCE_TO_LRT_METERS 4.588 4.99e-06
***SectorsEAST:SHORTEST_DISTANCE_TO_POLICE_METERS 2.662 0.007884**
SectorsSOUTH:SHORTEST_DISTANCE_TO_POLICE_METERS 1.964 0.049782* \
SectorsWEST:SHORTEST_DISTANCE_TO_POLICE_METERS -3.185 0.001489 \*\*
SectorsNORTH:male_percentage_num -3.817 0.000143
***SectorsNORTHEAST:male_percentage_num -2.897 0.003840**
SectorsSOUTH:male_percentage_num 1.982 0.047713* \
SectorsNORTHEAST:age_75_plus_percentage_num 2.505 0.012370 \*\
SectorsSOUTHEAST:age_75_plus_percentage_num 4.931 9.42e-07
***SectorsNORTH:TotalPermits -3.043 0.002396**
SectorsNORTHEAST:TotalPermits -4.637 3.95e-06*
**SectorsNORTHWEST:TotalPermits 3.595 0.000339**
*SectorsSOUTHEAST:TotalPermits -3.979 7.37e-05*
**SectorsNORTHWEST:median 4.215 2.70e-05** *SectorsSOUTH:median 5.711
1.44e-08* **SectorsSOUTHEAST:median 3.244 0.001214**
SHORTEST_DISTANCE_TO_LRT_METERS:SHORTEST_DISTANCE_TO_POLICE_METERS 2.026
0.043004 \*\
SHORTEST_DISTANCE_TO_LRT_METERS:male_percentage_num 3.571 0.000371
***SHORTEST_DISTANCE_TO_LRT_METERS:age_75_plus_percentage_num 3.063
0.002245** SHORTEST_DISTANCE_TO_LRT_METERS:TotalPermits 2.553 0.010821*
\
SHORTEST_DISTANCE_TO_POLICE_METERS:median -2.096 0.036288 \*\
male_percentage_num:age_75_plus_percentage_num -3.287 0.001045 \*\*
male_percentage_num:median -3.384 0.000740 \*\*
*median:canada_unemployment_rate_num 2.062 0.039430*

We will include these interaction terms in the model:

Year*male_percentage_num+\
Year*TotalPermits+\
Year*canada_unemployment_rate_num+\
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+\
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+\
Sectors*male_percentage_num+\
Sectors*age_75_plus_percentage_num+\
Sectors*TotalPermits+\
Sectors*median+\
SHORTEST_DISTANCE_TO_LRT_METERS*SHORTEST_DISTANCE_TO_POLICE_METERS+\
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+\
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+\
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+\
SHORTEST_DISTANCE_TO_POLICE_METERS*median+\
male_percentage_num*age_75_plus_percentage_num+\
male_percentage_num*median+\
median\*canada_unemployment_rate_num

Then we run the model with the interaction terms:

```{r}
model4 <- lm(ln_crime_rate ~ 
Year+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Year*canada_unemployment_rate_num+                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*SHORTEST_DISTANCE_TO_POLICE_METERS+   
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median+                                          
median*canada_unemployment_rate_num   
  , data_complete)


summary(model4)

```

Dropped the interaction terms had p\>0.05. They will be retained in the
model. median*canada_unemployment_rate_num\
Year*canada_unemployment_rate_num

```{r}

model4b <- lm(ln_crime_rate ~ 
Year+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*SHORTEST_DISTANCE_TO_POLICE_METERS+   
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete)


summary(model4b)

```

Next we move on to testing higher order terms. We use the function
pairs() to see how the response looks with respect to each independent
variable. We look at all pairwise combinations of continuous variables
in scatterplots.

```{r}
pairs(~ln_crime_rate +
Year+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete, panel=panel.smooth)



```

It looked like there might be some concavity in the

ln_crime_rate vs. Year, ln_crime_rate vs.SHORTEST_DISTANCE_TO_LRT_METERS
ln_crime_rate vs.SHORTEST_DISTANCE_TO_POLICE_METERS

We fitted a quadratic model and the linear model and see which one might
be the best fit.

```{r}

model5 <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete)

summary(model5)
```

```{r}
model5b <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+
I(Year^3)+    
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete)

summary(model5b)
```

```{r}

model5c <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+  
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete)

summary(model5c)

 
```

```{r}
model5c <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete)

summary(model5c)
```

I(SHORTEST_DISTANCE_TO_LRT_METERS\^4) not significant

```{r}
model5d <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^4)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete)

summary(model5d)
```

I(SHORTEST_DISTANCE_TO_POLICE_METERS\^2) NOT significant

```{r}
model5e <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+
SHORTEST_DISTANCE_TO_POLICE_METERS+
I(SHORTEST_DISTANCE_TO_POLICE_METERS^2)+  
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num, data_complete)

summary(model5e)
```

Then we fit a quadratic model plus the interaction terms.

```{r}

model6 <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*SHORTEST_DISTANCE_TO_POLICE_METERS+   
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete)

summary(model6)



```

```{r}

anova(model6, model4b)
AIC(model6)
AIC(model4b)

```

Dropped
interaction:SHORTEST_DISTANCE_TO_LRT_METERS\*SHORTEST_DISTANCE_TO_POLICE_METERS

```{r}
model6b <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete)

summary(model6b)
```

```{r}
anova(model6, model6b)
AIC(model6)
AIC(model6b)
```

The higher order term and the interaction terms were all significant at
alpha=0.05. This would be the final addictive model. We then run a
F-test for comparing model2 vs model6b.

$$
H_0:\\
H_a:\\
\alpha=0.05
$$

```{r}

anova (model2, model6b)

```

The $F_{cal}$=14.457, p \< 2.2e-16, reject $H_0$ that there was no
difference between model2 and model 6

Do we need to report any ANOVA Table ?

```{r}

df <- data.frame(
  Source_of_variation = c("Regression", "Residual", "Total"),
  DF = c(53, 1152, 1205),
  Sum_of_square = c(177.31, 220.51, 177.31 + 220.51),
  Mean_square = c(round(177.31/53,2), round(220.51/1152,2), ""),
  F = c(14.457, " ", " "),
  P_value = c("<2.2e-16", " ", " ")
)

kable(df, format = "html", align = "c", escape = FALSE) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(" " = 1, "ANOVA Table" = 5))

```

Model6b would be the final additive model,

Next we will check model assumptions:

1.  Linearity Assumption - Review residual plots (histogram and residual
    vs. fitted values)

The residual plots shows no discernible pattern.

```{r}
plot(model6b, which=1)
hist(model6b$residual, breaks=25)

```

2.  Independence Assumption - Review residual against year (time), and
    residual against year spatial variable (sectors)

```{r}
plot(data_complete$Year, model6b$residual, 
     xlab="Year of assessment") ## no discernible pattern over years?

boxplot(model6b$residual ~ data_complete$Sectors, 
        xlab="Sector") ## no discernible pattern across sectors?

boxplot(model6b$residual ~ data_complete$Year, 
        xlab="Year") 



```

3.  Normality Assumption - Using Shapiro-Wilk normality test; and normal
    qq plot, and histogram of residual

$$
H_0:\\
H_a:\\
\alpha=0.05
$$

```{r}
hist(model6b$residual, breaks=25)

plot(model6b, which=2) ## deviation at the tails 

shapiro.test(residuals(model6b)) ## not met

```

4.  Equal Variance Assumption (heteroscedasticity) - Using Breusch-Pagan
    test residual vs. fitted value plots

$$
H_0:\\
H_a:\\
\alpha=0.05
$$

```{r}


plot(model6b, which=1)

plot(model6b, which=3) ## the scale location plot showed a bit the variation of residuals decreases as fitted value increased; the line was quite straight. But it did not pass the bp test. To consult instructor, do we still need to do any data manipulation ?


bptest(model6) ## did not pass

```

5.  Multicollinearity - Using variance inflation factors (VIF)

To confirm with instructor: when testing for multicollinearity, we don't
include interaction terms and squared terms, so the model used to test
for multicollinearity would be model2, in which there were no variables
with VIF\>5.

```{r}
# imcdiag(model6b, method="VIF")
imcdiag(model2, method="VIF")

```

6.  Outliers - check Cook's distance and leverage

```{r}
plot (model6b, which=4) ## there was no outliers identified by Cook's distance >0.5
plot (model6b, which=5)
plot (model6b, which=6)


## leverage points

lev=hatvalues(model6b)
p = length(coef(model6b)) # 69
n = nrow(data_complete) # 1221
outlier2p = lev[lev>(2*p/n)]
outlier3p = lev[lev>(3*p/n)]
print("h_I>2p/n, outliers are")
outlier3p

# outlier2p = 2*69/1221 = 0.1130221
# outlier3p = 3*69/1221 = 0.1695332

```

Try to remove outlier3p and re-run model

```{r}

data_complete$lev <- hatvalues(model6b) 

data_complete_out <- data_complete[data_complete$lev<=0.1695332,]
  
```

```{r}


model7 <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete_out)

summary(model7)

plot(model7, which=1)
plot(model7, which=2)
plot(model7, which=3) # improved
shapiro.test(residuals(model7)) ## still not pass
bptest(model7)  ## still not pass

```

We have done the log transformation but still did not pass the
shapiro.test and bp test. boxcox is unlikely to help. Is Boxcox
transformation is still required? is the transformation performed on the
crime rate prior to log. But neither is helpful to pass the shapiro.test
and bp test.

```{r}
bc=boxcox(model6b,lambda=seq(-5,5))
bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda ## 1.161616

```

```{r}
model8 <- lm((((ln_crime_rate^(1.161616))-1)/(1.161616)) ~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete)

summary(model8)
plot(model8, which=1)
plot(model8, which=2)
plot(model8, which=3)
shapiro.test(residuals(model8))
bptest(model8)

```

Try Boxcox on the raw crime rate

```{r}
model9 <- lm(crime_rate~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete)

summary(model9)
plot(model9, which=1)
plot(model9, which=2)
plot(model9, which=3)
shapiro.test(residuals(model9))
bptest(model9)

```

```{r}
bc=boxcox(model9,lambda=seq(-5,5))
bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda ## 0.05050505
```

```{r}
model8 <- lm((((crime_rate^(0.05050505))-1)/(0.05050505)) ~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete)


summary(model10)
plot(model10, which=1)
plot(model10, which=2)
plot(model10, which=3)
shapiro.test(residuals(model10))
bptest(model10)
```

```{r}
# review : 304 crime count =1.
```

```{r}

data_complete_rm304 <- data_complete[data_complete$Total_Crime_Count>1,]

```

```{r}
model11 <- lm(ln_crime_rate ~ 
Year+
I(Year^2)+  
Sectors+
SHORTEST_DISTANCE_TO_LRT_METERS+
I(SHORTEST_DISTANCE_TO_LRT_METERS^2)+
I(SHORTEST_DISTANCE_TO_LRT_METERS^3)+   
SHORTEST_DISTANCE_TO_POLICE_METERS+
male_percentage_num+
age_75_plus_percentage_num+
TotalPermits+
median+
canada_unemployment_rate_num+
Year*male_percentage_num+                                            
Year*TotalPermits+                                                   
Sectors*SHORTEST_DISTANCE_TO_LRT_METERS+                         
Sectors*SHORTEST_DISTANCE_TO_POLICE_METERS+                      
Sectors*male_percentage_num+                                    
Sectors*age_75_plus_percentage_num+                         
Sectors*TotalPermits+                                           
Sectors*median+                                             
SHORTEST_DISTANCE_TO_LRT_METERS*male_percentage_num+                  
SHORTEST_DISTANCE_TO_LRT_METERS*age_75_plus_percentage_num+           
SHORTEST_DISTANCE_TO_LRT_METERS*TotalPermits+                         
SHORTEST_DISTANCE_TO_POLICE_METERS*median+                           
male_percentage_num*age_75_plus_percentage_num+                      
male_percentage_num*median, data_complete_rm304)


summary(model11)
plot(model11, which=1)
plot(model11, which=2)
plot(model11, which=3)
shapiro.test(residuals(model11)) # not pass
bptest(model11) # not pass

```

\`\`\`
